---
title: "Seoul Bike Sharing Demand Analysis and Prediction"
author: "Ahmad Sadeed (asadeed2), Deepa Nemmili Veeravalli (deepan2), Rui Zou (ruizou4)"
date: '2022-07-22'
output:
  pdf_document: default
  html_document: 
    theme: readable
    toc: yes
urlcolor: cyan
---

***

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.alin = "center")
```

## Description of the data file

This data file contains count of public bikes rented at each hour in Seoul Bike Sharing System with the corresponding weather data and holidays information. It has 14 variables and 8760 observations. We are interested in using Rented.Bike.Count (a numeric variable) as our response variable and explore how other factors (3 categorical variables and several continuous numeric variables) affect the count of bikes rented at each hour. Among the other 13 variables which we plan to use as potential predictors, we know from intuition that some may have more importance than others, like temperature, humidity, wind speed, visibility, seasons, and holiday, etc. 

## Background information on the data set

The original data comes from [http://data.seoul.go.kr](http://data.seoul.go.kr). The holiday information comes from [SOUTH KOREA PUBLIC HOLIDAYS](http://publicholidays.go.kr). A clean version can be found at [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Seoul+Bike+Sharing+Demand). 

Attribute Information:

- Date : month/day/year 
- Rented Bike count - Count of bikes rented at each hour 
- Hour - Hour of the day 
- Temperature - Temperature in Celsius 
- Humidity - % 
- Windspeed - m/s 
- Visibility - 10m 
- Dew point temperature - Celsius 
- Solar radiation - MJ/m2 
- Rainfall - mm 
- Snowfall - cm 
- Seasons - Winter, Spring, Summer, Autumn 
- Holiday - Holiday, No holiday 
- Functional Day - Functional or Non-functional days of rental bike system

## Our Interest
This data set is interesting to us both personally and business-wise. Recently we have seen a rise in the delivery, accessibility, and usage of regular and electric rental bikes. There are clear environmental, health, and economical benefits associated with the usage of bikes as a mode of transportation. We would like to find out what factors lead to an increase in number of bikes rented and what factors have inverse effect on using rental bikes. Learning about such factors can help a bike rental business manage its inventory and supply without any hindrance. It can also help cities plan accordingly due to an increase of bikers, e.g. opening up more bike lanes during certain days or seasons. Environmentally, we will have a better understanding of the feasibility of turning a city into a "bike city" or looking at alternative options if a city is not friendly to bikers due to harsh weather conditions. 

## Data in R

The data file can be successfully loaded into R. We have printed out the structure and first few rows of the data file below.

The column names in the `csv` file contains measurement units (like `Wind speed (m/s), Solar Radiation (MJ/m2)`) and characters such as $^\circ$ and %. We load the data using cleaned up column names.  

```{r}
columns = c("Date","Rented","Hour","Temp","Humidity","Wind","Visibility","Dew",
            "Radiation","Rain","Snow","Season","Holiday","Functioning")
bike = read.csv("../data/SeoulBikeData.csv", col.names = columns)
str(bike)
```

```{r message=FALSE, warning=FALSE}
library(lubridate)

bike$Date = as.Date(bike$Date, '%d/%m/%Y')
bike$Weekday = weekdays(bike$Date, abbreviate=TRUE)
# bike$Weekend = ifelse(bike$Weekday == "Sun" | bike$Weekday == "Sat", "Yes", "No")

# bike$Hour = as.factor(bike$Hour)
bike$Season = as.factor(bike$Season)
bike$Holiday = as.factor(bike$Holiday)
bike$Functioning = as.factor(bike$Functioning)
bike$Weekday = as.factor(bike$Weekday)

str(bike)
```

The data is for one year (from Dec 2017 to Dec 2018)

```{r}
library(ggplot2)
ggplot(data = bike, aes(x = Date, y = Rented)) +
    scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
    theme(axis.text.x=element_text(angle=60, hjust=1)) +
    geom_bar(stat = "identity", fill = "blue") +
    labs(title = "Number of bikes rented ",
         subtitle = "2017 December to November 2018",
         x = "Date", y = "Rented Bikes Count")
```


```{r message=FALSE, warning=FALSE}
library(corrr)
library(dplyr)
bike_num = subset(bike, select = -c(Date, Hour, Season, Holiday, Functioning, Weekday) )
correlations = corrr::correlate(bike_num)
top_3 = head(dplyr::arrange(corrr::stretch(correlations, remove.dups = TRUE), desc(r)), 5)
top_3
```

Drop Data and Dew:

```{r}
bike_cln= subset(bike, select = -c(Date, Dew) )
str(bike_cln)
```

```{r}
mod_additive = lm(Rented ~ ., data = bike_cln)
summary(mod_additive)$r.squared
summary(mod_additive)$adj.r.squared
```

Checking for multicollinearity in the data, we see that `Temp` has a rather high variance inflation factor. 

```{r message=FALSE, warning=FALSE}
library(faraway)
vif(mod_additive)
```

```{r}
# copy dataset to test Hour as factor variable
bike_hr = data.frame(bike_cln)
bike_hr$Hour = as.factor(bike_hr$Hour)
# _hr for using Hour as factor
mod_additive_hr = lm(Rented ~ ., data = bike_hr)
summary(mod_additive_hr)$r.squared
summary(mod_additive_hr)$adj.r.squared
```


```{r}
vif(mod_additive_hr)
```

```{r}
mod_interact = lm(Rented ~ .^2, data = bike_cln)
summary(mod_interact)$r.squared
summary(mod_interact)$adj.r.squared
```

```{r}
mod_interact_hr = lm(Rented ~ .^2, data = bike_hr)
summary(mod_interact_hr)$r.squared
summary(mod_interact_hr)$adj.r.squared
```

```{r}
anova(mod_additive, mod_additive_hr)
anova(mod_interact, mod_interact_hr)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr) 
library(tibble) 
library(dlookr)
```


```{r}
diagnose_numeric(bike_cln)
```


```{r}
bike_cln %>% plot_outlier(Rented)
```

This will generate an HTML report of the data:

```{r message=FALSE, warning=FALSE, include=FALSE}
# diagnose_web_report(bike_cln)
```


```{r}
normality(bike_cln)
```

Check normality of each numeric variable and 

```{r}
bike_hr %>% plot_normality(Rented, Temp, Humidity, Wind, Visibility, Radiation,
                           Rain, Snow)
```

```{r}
# library(caret)
# cc = caret::preProcess(data.frame(bike_cln$Temp), method=c("BoxCox"))
# bike_cln$Temp = predict(cc, bike_cln$Temp)
```

We will use Hour as factor going forward.

This will generate an HTML report of the data:

```{r include=FALSE}
# install.packages('forecast', dependencies = TRUE)
# eda_web_report(bike_hr)
```

Check model after taking sqrt of Rented:

```{r}
# add 1 to response variable to avoid errors in log
bike_hr$Rented = bike_hr$Rented + 1
```

```{r}
# mod_interact_sq = lm(Rented ~ .^2, data = bike_sqrt)
mod_interact_sq = lm(sqrt(Rented) ~ .^2, data = bike_hr)
summary(mod_interact_sq)$r.squared
summary(mod_interact_sq)$adj.r.squared
```

Check log transformation on Rented:

```{r}
# mod_interact_log = lm(Rented ~ .^2, data = bike_log)
mod_interact_log = lm(log(Rented) ~ .^2, data = bike_hr)
summary(mod_interact_log)$r.squared
summary(mod_interact_log)$adj.r.squared
```

This model was picked after running AIC on the model with all interactions:

```{r}

# mod_int_aic = step(mod_interact, direction = "backward", trace = 0)
mod_int_aic = lm(formula = Rented ~ Hour + Temp + Humidity + Wind + Visibility + 
    Radiation + Rain + Snow + Season + Holiday + Functioning + 
    Weekday + Hour:Temp + Hour:Humidity + Hour:Visibility + Hour:Radiation + 
    Hour:Rain + Hour:Season + Hour:Holiday + Hour:Functioning + 
    Hour:Weekday + Temp:Humidity + Temp:Wind + Temp:Radiation + 
    Temp:Snow + Temp:Season + Temp:Holiday + Temp:Functioning + 
    Temp:Weekday + Humidity:Wind + Humidity:Visibility + Humidity:Radiation + 
    Humidity:Rain + Humidity:Season + Humidity:Functioning + 
    Humidity:Weekday + Wind:Visibility + Wind:Radiation + Wind:Rain + 
    Wind:Season + Wind:Weekday + Visibility:Radiation + Visibility:Snow + 
    Visibility:Season + Visibility:Holiday + Visibility:Weekday + 
    Radiation:Season + Radiation:Holiday + Radiation:Functioning + 
    Radiation:Weekday + Rain:Season + Rain:Functioning + Rain:Weekday + 
    Snow:Season + Season:Holiday + Season:Functioning + Season:Weekday + 
    Holiday:Functioning + Holiday:Weekday, data = bike_cln)

summary(mod_int_aic)$r.squared
summary(mod_int_aic)$adj.r.squared
```

Check cross-validated RMSE of different models:

```{r}
calc_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}
```


```{r}
calc_loocv_rmse(mod_additive)
calc_loocv_rmse(mod_additive_hr)
calc_loocv_rmse(mod_interact)
calc_loocv_rmse(mod_interact_hr)
calc_loocv_rmse(mod_int_aic)

# can't use this fuction to compare models with log(Rented) or sqrt(Rented)
# calc_loocv_rmse(mod_interact_sq) # the output doesn't mean the same here
# calc_loocv_rmse(mod_interact_log)
```

The model with log transformation of `Rented` achieves the best LOOCV RMSE.

Split data into train/test to test the models:

```{r}
set.seed(420)
bike_idx = sample(1:nrow(bike_hr), 8000)
bike_trn = bike_hr[bike_idx, ]
bike_tst = bike_hr[-bike_idx, ]
```

```{r}
RMSE <- function(model, data, trans = "") {
  n = nrow(data)
  y_hat = predict(model, data)
  if(trans=="log") {
      resid = data$Rented - exp(y_hat)
  } else if (trans=="sqrt"){
      resid = data$Rented - y_hat^2
  } else {
      resid = data$Rented - y_hat
  }
  sqrt(sum(resid ^ 2) / n)
}
```


```{r warning=FALSE}
mod_additive_trn = lm(Rented ~ ., data = bike_trn)
RMSE(mod_additive_trn, bike_tst)

mod_interact_trn = lm(Rented ~ .^2, data = bike_trn)
RMSE(mod_interact_trn, bike_tst)

mod_interact_sq_trn = lm(sqrt(Rented) ~ .^2, data = bike_trn)
RMSE(mod_interact_sq_trn, bike_tst, trans = "sqrt")

mod_interact_log_trn = lm(log(Rented) ~ .^2, data = bike_trn)
RMSE(mod_interact_log_trn, bike_tst, trans = "log")

```

Check fitted vs residuals for some models:

```{r}
par(mfrow = c(1, 2))

plot(fitted(mod_interact_trn), resid(mod_interact_trn), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(mod_interact_trn), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(mod_interact_trn), col = "dodgerblue", lwd = 2)
```

```{r}
par(mfrow = c(1, 2))

plot(fitted(mod_interact_sq_trn), resid(mod_interact_sq_trn), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(mod_interact_sq_trn), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(mod_interact_sq_trn), col = "dodgerblue", lwd = 2)
```



```{r}
par(mfrow = c(1, 2))

plot(fitted(mod_interact_log_trn), resid(mod_interact_log_trn), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(mod_interact_log_trn), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(mod_interact_log_trn), col = "dodgerblue", lwd = 2)
```






